Implementation of a Brute Force Attack on the A5/1 Keystream Generator in a GPU-based Volunteer Computing Project

### 2017-02-05 Paper PaCT 2017/pact2017_bulavintsev_rc2.tex
fixed the project decription and Related work

\title{Implementation of a Brute Force Attack on A5/1 Cryptographical
Algorithm in a GPU-based Volunteering Computing Project}
%If Title is too long, use \titlerunning
\titlerunning{A5/1 Brute Force on GPU}

%Single insitute
\author{Bulavintsev V. \and Semenov A. \and Zaikin O.}
%If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}
\institute{Matrosov Institute for System Dynamics and Control Theory of
Siberian Branch of Russian Academy of Sciences
(ISDCT SB RAS)}

@@

\begin{abstract}
We present an advanced brute force attack on the A5/1 cryptoalgorithm,
	that is still broadly used in modern GSM networks. We use a
	well-known idea introduced by R.Anderson more than 20 years ago
	to greatly reduce the search space. The primary contribution of
	this article is the implementation of the Anderson's attack on a GPU platform
	with the bitslice technique. The preliminary estimates of the
	attack's speed showed that, with the use of a GPUs processing power, 
	the attack could be conducted in the real
	time on a modern computer cluster or in a volunteer computing
	project. To verify our estimates with the use of BOINC technology 
	we concieved the specialized volunteering computing project 
	and executed our variant of Anderson's attack within it. In the
	7 days the project lasted, 10 A5/1 cryptanalysis problems were
	solved. The results presented in this work provide yet another
	proof of A5/1's cryptographical weakness that make it totally
	unsuitable for transmission of any kind of sensitive data
	through modern GSM networks.

@@

\begin{keywords}
A5/1 algorithm, brute-force cryptanalysis, GPU, volunteering computing
\end{keywords}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{./a51.png} 
	\caption{A5/1 generator work scheme} 
	\label{fig:a51gen} 
\end{figure}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{./a51and.png} 
	\caption{The set of guessing bits used in Anderson's attack (greyed out).}
	\label{fig:a51and}
\end{figure}

@@

The effeciency of a brute force attack is defined by two parameters: the speed
of checking of the key candidates and the size of the search space. R.
Anderson's idea described above give us a $$2^{53}$$ search space for the A5/1
generator. Instead of using the 'naive' implementation of A5/1 generator, to
speed up the key candidates checking procedure one can opt to use more
sophisticated alternatives. We evaluated the performance of two different fast
implementations of A5/1 generator in \cite{PAVT_2016}. The first one was based
on an idea of precomputation of the states of the registers LFSR 1-3, and
keeping the states in RAM, to use them as the look-up tables. This approach
demonstrated a considerable speed-up against 'naive' implementation. A similiar
method of precomputation of LFSRs was described in \cite{BSW}. However, in
\cite{PAVT_2016} we found its performance inferior to the second
implementation, that is based on the 'bit-slicing technique'. Next we briefly
describe the idea of this technique, and it's application to Anderson's attack.

Modern general-purpose computational architectures are suboptimal for
implementation of many algorithms. The last operate with bits, while CPUs and
GPUs typically operate with 32-256 bit words. As the result of this
discrepancy, the 'naive' implementation of a cryptographical algorithm won't be
fully using the platform's computational resources. For example, to conduct a
'addition modulo 2' (XOR) operation on 2 boolean arguments, a modern CPU will
position the arguments values into the lowest bits of two 32-bit
general-purpose register (GPR), perform the computation and write the result
into the third 32-bit GPR. In effect, 31 out of 32 bits of GPRs stay idle
during this operation. This problem could be solved by using bitwise operations
that work on GPRs like on boolean vectors - digit by digit. A cryptographical
algorithm could be oftentimes represented as a logical scheme built with the
basic logical gates. Coupled with bitwise operations, this representation
allows the computational platform to apply the algorithm to as many sets of
data, as there are bits in it's GPR (e.g. the platform's register capacity).
This method is called 'SIMD within a register' or the 'bit-slicing' technique.
The first mention of bit-slicing technique applied to cryptographical problem
belongs, seemingly, to E.Biham \cite{BIT_SL}.

@@

\begin{figure} \includegraphics[width=\linewidth]{./bitslice.png}
	\caption{Simultaneous computation of $D$ instances of scheme $C(f)$ with
	bit-slicing technique} \label{fig:a51gen} 
\end{figure}

@@

The Anderson's attack algorithm described above was realized on an NVIDIA GPU
with the use of CUDA SDK 8.0 \cite{CUDA}. The comparison of the performance of
a GPU to a CPU in execution of bit-slicing and LFSR precomputation-based
\cite{BSW} implementations of Anderson's attack is shown in
\cite{tab:gpuspeed}.

\begin{table} \caption{(LOP3.LUT is a special instruction that implements
	arbitrary bitwise arity 3 functions in hardware. We use it as a substitute
	for the "bitselect" instruction.)} 
	\label{tab:gpuspeed} Performance of the different implementations of the
	Anderson's attack (search space is $2^{53}$) on a CPU and a GPU, measured
	in millions of key checks per second.  
	\begin {center} 
	\begin{tabular} {| l | l | l |} \hline Computational device & Bit-slicing &
		LFSR precomputation \cite{BSW} \\ \hline CPU Intel Core I7 930 & 37 & 7
		\\ \hline GPU NVIDIA GTX 1050 Ti & 9180 & 483 \\ \hline GPU NVIDIA GTX
		1050 Ti (LOP3.LUT) & 11950 & - \\ \hline
	\end{tabular}
\end {center}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation of the Anderson's Attack in a volunteer computing
project}\label{sec:boinc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In order to solve 10 cryptanalysis problems for the A5/1 keystream generator we
launched the volunteer computing project AndersonAttack@home. The computing
application of this project is based on the CUDA implementation, which was
described in the previous section.

Volunteer computing is a type of distributed computing which uses computational
resources of PCs (hosts) of private persons called volunteers. Each volunteer
computing project is designed to solve one or several hard problems. When PC is
connected to the project, all the calculations are performed automatically and
do not inconvenience user since only idle resources of PC are used. Volunteers
get credits (points) for each calculated task.

Volunteer computing project consists of the following basic parts: server
daemons, database, web site and client application. Daemons include
\textsc{work generator} (generates tasks to be processed), \textsc{validator}
(checks the correctness of the results received from volunteer's PCs) and
\textsc{assimilator} (processes correct results). Client application should
have versions for the widespread computing platforms.  One of attractive
features of volunteer computing is its low cost --- to maintain a project one
only needs a dedicated server working 24/7.

AndersonAttack@home is based on BOINC (Berkley Open Infrastructure for Network
Compuitng \cite{DANDERSON2004}), which is the most popular platform for
volunteer computing. Each BOINC-based project is separate, but any private
person can take part in arbitrary project with the help of standard BOINC
manager. This BOINC manager allows users to change projects priority and to
regulate the mode of PC using.

BOINC provides a form of redundant computing.  According to this, several (at
least 2) similar tasks are created for each workunit. These tasks are processed
on multiple hosts, the results are compared on a project server by validator,
and are accepted only when a 'consensus' is reached. The sufficient number of
successful results is called a 'quorum'. If for a particular workunit results
of tasks are nonsimilar, then new tasks must be created and sent. The scheme of
a BOINC-based project with the quorum of 2 is shown on the Fig. 1.

\begin{figure} \includegraphics[width=\linewidth]{./boinc.png}
	\caption{Simultaneous computation of $D$ instances of scheme $C(f)$ with
	bit-slicing technique} \label{fig:a51gen} 
\end{figure}

On the first stage of our experiment on AndersonAttack@home, a family of
workunits was generated on the project server. In each workunit values of 12
variable cells out of 53 (see the previous section) were fixed.  Thus, 40960
workunits were generated for 10 cryptanalysis instance in total. Usually the
value of deadline for workunits in BOINC projects is equal to 10-14 days. In
our project we used the deadline of 1 day, because the experiment was quite
small.  On the next stage all generated workunits were processed in a desktop
grid formed by project’s hosts. This took about 7 days. As a result, solutions
for all considered instances were successfully found. All these solutions are
listed in Table 1. It should be noted, that collisions were found for 7 out of
10 instances (5 of them have 1 collision, each of other 2 instances has 2
collisions).

\begin{table} \caption{} \label{tab:keys} Original secret keys and collisions of the
	generator A5/1 (in hexadecimal format) which were found in
	AndersonAttack@home.
	\begin {center} 
		\begin{tabular}
			{| l | l | l |}
			\hline
			Instance & Keystream & Secret key \\ \hline
			1 & 0x770c0410869366f1 & original 0x11b8e4340276c4ee \\
			  &                    & collision1 0x42634f3266d302a3 \\ \hline
			2 & 0xae9590560c26e9ed & original 0x4c656fd73e59ab9b \\
			  &                    & collision1 0xcf23e4722e3cfb68 \\ \hline
			3 & 0xdd4b3ab7f6cf8224 & original 0x09429d158555f4b3 \\
			  &                    & collision1 0x09429d158553e967 \\
			  &                    & collision2 0x40e5f2c8128a1781 \\ \hline
			4 & 0x93cd42d97eb75fd9 & original 0xfa386a338355aafd \\
			  &                    & collision1 0xf9e81096bb4d0aad \\ 
			  &                    & collision2 0xf9e81096bb4a8556 \\ \hline
			5 & 0x925e423c98121152 & original 0xe5cf81035ce5fbe2 \\ \hline
			6 & 0x3b3464bd6e377b87 & original 0x9625e9d810b46248 \\
			  &                    & collision1 0xf5aa1be2d6c36e18 \\ \hline
			7 & 0x0367d29121dd1677 & original 0xd1b8b06086edf162 \\ \hline
			8 & 0x6b49230b7fc0249d & original 0xbe81a896968c486b \\ \hline
			9 & 0xc65847556752d14c & original 0xb6f65d2855a211c0 \\ 
			  &                    & collision1 0xb6f65d2855a508e0 \\ \hline
			10& 0x07bb7f83d26072ec & original 0x122a1a2955286b9f \\ 
			  &                    & collision1 0xd5151aaa50490012 \\ \hline
		\end{tabular}
	\end {center}
\end{table}

According to BOINC statistics, 143 active hosts of 90 volunteers participated
in the experiment.  Here by ‘active’ we mean a host which correctly processed
at least one task.  Active volunteer has at least one active host. It should be
noted, there is only 1 volunteer from Russia in top-10 project volunteers list
(formed according to the amount of credits), the most of them are from United
States (4 persons). In top-10 list of teams there is only one team from Russia
too.

@@

As was noted before, at the time of writing of our article the A5/1 algorithm
remains one of the most popular research subjects among cryptography
specialists, along with DES and RSA algorithms. The exact authorship of this
algorithm is unknown, but some experts argue that chances are high that it's
creation should be attributed to a French special agencies. Numerous sources
describe the events that led to it's structure leaked to general public. In
fact, it's internal workings were already known to research community in 1994:
R.Anderson's note (a mailing list message) written in that period was one of
the firs works on A5/1 cryptanalysis. The complete knowledge of the A5/1 were
acquired in 1999 as a result of a reverse engineering of a mobile phone.

Earlier we mentioned Anderson's attack (described by Ron Anderson in a short
note in 1994 \cite{ANDERSON}) was the first attack on A5/1 with compexity less
than that of a trivial brute-force walk over the whole keyspace. Shortly
afterwards J.Golic suggested the "Alleged" attack on the A5/1, based on a
linearization of the equations describing A5/1 \cite{GOLIC}. The Golic's attack
estimated complexity is $C\dot2^{40}$, where $C$ denoting the complexity of
solving the system of linear equations over $GF(2)$ with a very substanial
number of dimensions. The next widely known attack on the A5/1 was presented in
\cite{BSW}. To speed up the A5/1 algorithm it's authors used the technique of
precomputation of LFSRs we mentioned in \cref{sec:bitsl}. It's worth to note
that the attack presented in \cite{BSW}, along with those presented in
\cite{BIKHAM2000}, \cite{EKDAHL2003}, \cite{BARKAN2008} \todo{Check the actual
articles! Not sure if these point to the right ones.}, require substanial
lengths of known keystream (several seconds at best). These attacks can be
deemed realistic, but they do not demonstrate vulnerability of the GSM security
protocols as evidently, as those belonging to the "advanced brute force" class,
that we will discuss below. For the rest of the section we consider an attack
practical if it's real-life implementation is able to persistently find the
secret key by analyzing no more than 2 frames (912 bits) of the keystream, the
amount that could always be extracted from a call in a GSM network by
exploiting the protocol vulnerability described in \Cref{sec:alg}.

It seems that the first practical attack ('practical' in a sense we established
above) was presented in \cite{COPAC_1}. It's authors implemented the optimezed
variant of the Anderson's attack on specialized computational device of their
own design, assembled from 120 "Xilinx Spartan 3" FPGAs. The authors state
\cite{COPAC_1} that the attack took about 6 hours to complete. It's worth
mentioning that the COPACOBANA architecture was used for cryptanalytical
research of several other ciphers, e.g. DES \cite{COPAC_2}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{splncs03}
\bibliography{paper}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\title{Implementation of a Brute Force Attack on the A5/1 Keystream Generator
in a GPU-based Volunteer Computing Project}
\titlerunning{Brute Force Attack on A5/1}

%Single insitute
\author{Vadim Bulavintsev \and Alexander Semenov \and Oleg Zaikin}
\authorrunning{Vadim Bulavintsev et al.}
\institute{Matrosov Institute for System Dynamics and Control Theory of
Siberian Branch of Russian Academy of Sciences, Irkutsk, Russia
\email{ichorid@mail.ru, biclop.rambler@yandex.ru, zaikin.icc@gmail.com}}

@@

\begin{abstract}
We present an advanced brute force attack on the A5/1 keystream generator,
	that is still widely used in modern GSM networks. We use a
	well-known idea introduced by R. Anderson more than 20 years ago
	to greatly reduce the search space. The main contribution of
	the present paper is the implementation of the Anderson's attack on a GPU platform
	with the bitslice technique. The preliminary estimates of the
	attack's speed showed that, with the use of GPUs processing power, 
	the attack could be performed in the real time on a modern computer cluster or in a volunteer computing
	project. To verify our estimates with the use of the BOINC technology 
	we launched the specialized volunteer computing project 
	and executed our variant of Anderson's attack within it. 10 A5/1 cryptanalysis problems were
	solved in 7 days in the project. The results presented in this work provide yet another
	proof of A5/1's cryptographical weakness that make it totally
	unsuitable for transmission of any kind of sensitive data
	through modern GSM networks.

@@

\begin{keywords}
keystream generator, A5/1, cryptanalysis, brute force attack, GPU, volunteer computing, BOINC
\end{keywords}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{./a51.eps} 
	\caption{A5/1 generator work scheme} 
	\label{fig:a51gen} 
\end{figure}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{./a51and.eps} 
	\caption{The set of guessing bits used in Anderson's attack (greyed out).}
	\label{fig:a51and}
\end{figure}

@@

The effeciency of a brute force attack is defined by two parameters: the speed
of checking of the key candidates and the size of the search space. R.
Anderson's idea described above give us a $$2^{53}$$ search space for the A5/1
generator. Instead of using the 'naive' implementation of A5/1 generator, to
speed up the key candidates checking procedure one can opt to use more
sophisticated alternatives. We evaluated the performance of two different fast
implementations of A5/1 generator in \cite{PAVT_2016}. The first one was based
on an idea of precomputation of the states of the registers LFSR 1-3, and
keeping the states in RAM, to use them as the look-up tables. This approach
demonstrated a considerable speed-up against 'naive' implementation. A similiar
method of precomputation of LFSRs was described in \cite{DBLP:conf/fse/BiryukovSW00}. However, in
\cite{PAVT_2016} we found its performance inferior to the second
implementation, that is based on the 'bit-slicing technique'. Next we briefly
describe the idea of this technique, and it's application to Anderson's attack.

Modern general-purpose computational architectures are suboptimal for
implementation of many algorithms. The last operate with bits, while CPUs and
GPUs typically operate with 32-256 bit words. As the result of this
discrepancy, the 'naive' implementation of a cryptographical algorithm won't be
fully using the platform's computational resources. For example, to conduct a
'addition modulo 2' (XOR) operation on 2 boolean arguments, a modern CPU will
position the arguments values into the lowest bits of two 32-bit
general-purpose register (GPR), perform the computation and write the result
into the third 32-bit GPR. In effect, 31 out of 32 bits of GPRs stay idle
during this operation. This problem could be solved by using bitwise operations
that work on GPRs like on boolean vectors - digit by digit. A cryptographical
algorithm could be oftentimes represented as a logical scheme built with the
basic logical gates. Coupled with bitwise operations, this representation
allows the computational platform to apply the algorithm to as many sets of
data, as there are bits in it's GPR (e.g. the platform's register capacity).
This method is called 'SIMD within a register' or the 'bit-slicing' technique.
The first mention of bit-slicing technique applied to cryptographical problem
belongs, seemingly, to E. Biham \cite{BIT_SL}.

@@

\begin{figure} \includegraphics[width=\linewidth]{./bitslice.eps}
	\caption{Simultaneous computation of $D$ instances of scheme $C(f)$ with
	bit-slicing technique} \label{fig:a51gen} 
\end{figure}

@@

The Anderson's attack algorithm described above was realized on an NVIDIA GPU
with the use of CUDA SDK 8.0 \cite{CUDA}. The comparison of the performance of
a GPU to a CPU in execution of bit-slicing and LFSR precomputation-based
\cite{DBLP:conf/fse/BiryukovSW00} implementations of Anderson's attack is shown in
\cite{tab:gpuspeed}.

\begin{table} \caption{(LOP3.LUT is a special instruction that implements
	arbitrary bitwise arity 3 functions in hardware. We use it as a substitute
	for the "bitselect" instruction.)} 
	\label{tab:gpuspeed} Performance of the different implementations of the
	Anderson's attack (search space is $2^{53}$) on a CPU and a GPU, measured
	in millions of key checks per second.  
	\begin {center} 
	\begin{tabular} {| l | l | l |} \hline Computational device & Bit-slicing &
		LFSR precomputation \cite{DBLP:conf/fse/BiryukovSW00} \\ \hline CPU Intel Core I7 930 & 37 & 7
		\\ \hline GPU NVIDIA GTX 1050 Ti & 9180 & 483 \\ \hline GPU NVIDIA GTX
		1050 Ti (LOP3.LUT) & 11950 & - \\ \hline
	\end{tabular}
\end {center}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation of the Anderson's Attack in a volunteer computing
project}\label{sec:boinc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In order to solve 10 cryptanalysis problems for the A5/1 keystream generator we
launched the volunteer computing project AndersonAttack@home. The client (computing)
application of this project is based on the CUDA implementation, which was
described in the previous section.

Volunteer computing \cite{DBLP:conf/ccgrid/AndersonF06} is a type of distributed computing \cite{Foster:1995:DBP:527029} which uses computational resources of PCs (hosts) of private persons called volunteers. 
Each volunteer computing project is designed to solve one or several hard scientific problems. When PC is
connected to a project, all calculations are performed automatically and
do not inconvenience user since only idle PC resources are used.

Each volunteer computing project consists of the following basic parts: server
daemons, database, web site and client application. Daemons include
\textsc{work generator} (it generates tasks to be processed), \textsc{validator}
(it checks the correctness of the results received from volunteer's PCs) and
\textsc{assimilator} (aimed at processing correct results). A client application should
have versions for the widespread computing platforms.  One of attractive
features of volunteer computing is its low cost --- to maintain a project one
only needs a dedicated server working 24/7.

AndersonAttack@home is based on BOINC (Berkley Open Infrastructure for Network
Compuitng \cite{Anderson:2004:BSP:1032646.1033223}), which is the most popular platform for
volunteer computing. Each BOINC-based project is separate, but any private
person can take part in arbitrary project with the help of the standard BOINC
manager. This manager allows users to change projects priority and to
regulate the mode of PC using.

BOINC provides a form of redundant computing.  According to this, several (at
least 2) similar tasks are created for each workunit. These tasks are processed
on multiple hosts, the results are compared on a project server by validator,
and are accepted only when a 'consensus' is reached. The sufficient number of
successful results is called a 'quorum'. If for a particular workunit results
of tasks are nonsimilar, then new tasks must be created and sent. The scheme of
a BOINC-based project with the quorum of 2 is shown on Fig.~\ref{fig:boinc}.

%\begin{figure}
%\includegraphics[width=\linewidth]{./BOINC_project_scheme.eps}
%	\caption{The scheme of a BOINC-based project with the quorum of 2} 
%	\label{fig:boinc} 
%\end{figure}

In the first stage of our experiment, a family of workunits was generated on 
the project server. In each workunit values of 12 out of 53 cells of the A5/1 generator (see the previous section) 
were fixed. Thus, 40960 workunits were generated for 10 cryptanalysis problems in total. Usually, the
value of deadline for workunits in BOINC projects is equal to 10-14 days. In
our project, we used a deadline of 1 day, because the experiment was quite
small. In the next stage all generated workunits were processed in a desktop
grid formed by project's hosts. This took about 7 days. As a result, solutions
for all considered problems were successfully found (see Table~\ref{tab:keys}). It should be noted, 
that we found collisions for 7 out of 10 problems (5 of them have 1 collision, each of other 2 problems has 2
collisions).

\begin{table} 
\caption{} 
\label{tab:keys} Original secret keys and collisions of the
	generator A5/1 (in hexadecimal format) which were found in
	AndersonAttack@home.
	\begin {center} 
		\begin{tabular}
			{| l | l | l |}
			\hline
			Instance & Keystream & Secret key \\ \hline
			1 & 0x770c0410869366f1 & original 0x11b8e4340276c4ee \\
			  &                    & collision 1 0x42634f3266d302a3 \\ \hline
			2 & 0xae9590560c26e9ed & original 0x4c656fd73e59ab9b \\
			  &                    & collision 1 0xcf23e4722e3cfb68 \\ \hline
			3 & 0xdd4b3ab7f6cf8224 & original 0x09429d158555f4b3 \\
			  &                    & collision 1 0x09429d158553e967 \\
			  &                    & collision 2 0x40e5f2c8128a1781 \\ \hline
			4 & 0x93cd42d97eb75fd9 & original 0xfa386a338355aafd \\
			  &                    & collision 1 0xf9e81096bb4d0aad \\ 
			  &                    & collision 2 0xf9e81096bb4a8556 \\ \hline
			5 & 0x925e423c98121152 & original 0xe5cf81035ce5fbe2 \\ \hline
			6 & 0x3b3464bd6e377b87 & original 0x9625e9d810b46248 \\
			  &                    & collision 1 0xf5aa1be2d6c36e18 \\ \hline
			7 & 0x0367d29121dd1677 & original 0xd1b8b06086edf162 \\ \hline
			8 & 0x6b49230b7fc0249d & original 0xbe81a896968c486b \\ \hline
			9 & 0xc65847556752d14c & original 0xb6f65d2855a211c0 \\ 
			  &                    & collision 1 0xb6f65d2855a508e0 \\ \hline
			10& 0x07bb7f83d26072ec & original 0x122a1a2955286b9f \\ 
			  &                    & collision 1 0xd5151aaa50490012 \\ \hline
		\end{tabular}
	\end {center}
\end{table}

According to BOINC statistics, 143 active hosts of 90 volunteers participated
in the experiment. Here by 'active' we mean a host which correctly processed
at least one task. Active volunteer has at least one active host. It should be
noted, that there is only 1 volunteer from Russia in top-10 project's volunteers list
(formed according to the amount of calculations), the most of them are from United
States (4 persons). In top-10 list of teams there is only one team from Russia
too.

@@

As was noted before, at the time of writing of our article the A5/1 algorithm
remains one of the most popular research subjects among cryptography
specialists, along with DES and RSA algorithms. The exact authorship of this
algorithm is unknown, but some experts argue that chances are high that its
creation should be attributed to a French special agencies. Numerous sources
describe the events that led to its structure leaking to general public. In
fact, its internal workings were already known to research community in 1994:
R. Anderson's note (a mailing list message) written in that period was one of
the first works on A5/1 cryptanalysis. The complete knowledge of A5/1 were
acquired in 1999 as a result of a reverse engineering of a mobile phone.

Earlier we mentioned Anderson's attack (described by Ron Anderson in a short
note in 1994 \cite{ANDERSON}) was the first attack on A5/1 with compexity less
than that of a trivial brute-force walk over the whole keyspace. Shortly
afterwards J. Golic suggested the ``Alleged'' attack on the A5/1, based on a
linearization of the equations describing A5/1 \cite{Golic1997}. The Golic's attack
estimated complexity is $C \cdot 2^{40}$, where $C$ denotes the complexity of
solving the system of linear equations over $GF(2)$ with a very substanial
number of dimensions. The next widely known attack on the A5/1 was presented in
\cite{DBLP:conf/fse/BiryukovSW00}. To speed up the A5/1 algorithm its authors used the technique of
precomputation of LFSRs we mentioned in \Cref{sec:bitslc}. It's worth to note
that the attack presented in \cite{DBLP:conf/fse/BiryukovSW00}, along with those presented in
\cite{Biham2000,DBLP:journals/tit/EkdahlJ03,Barkan2008}, require substanial
lengths of known keystream (several seconds at best). These attacks can be
deemed realistic, but they do not demonstrate vulnerability of the GSM security
protocols as evidently, as those belonging to the ``advanced brute force'' class,
that we will discuss below. For the rest of the section we consider an attack
practical if its real-life implementation is able to persistently find the
secret key by analyzing no more than 2 frames (912 bits) of a keystream, the
amount that could always be extracted from a call in a GSM network by
exploiting the protocol vulnerability described in \Cref{sec:alg}.

It seems that the first practical attack (``practical'' in a sense we established
above) was presented in \cite{Guneysu:2008:CC:1446228.1446266}. It's authors implemented the optimized
variant of the Anderson's attack on a specialized computational device of their
own design, assembled from 120 ``Xilinx Spartan 3'' FPGAs. The authors state
\cite{Guneysu:2008:CC:1446228.1446266} that the attack took about 6 hours to complete. It's worth
mentioning that the COPACOBANA architecture was used for cryptanalytical
research of several other ciphers, e.g. DES \cite{COPAC_2}.

@@

\section*{Acknowledgment}
The research was funded by Russian Science Foundation (project No. 16-11-10046). The authors would like to thank Stepan Kochemazov for helpful discussions and valuable comments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{splncs03}
\bibliography{pact2017_bulavintsev_ref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
